Return Oriented Programming

### 2022-06-19 RetOriProg_Maximilian_Heim.tex
changes

\documentclass[11pt]{beamer}
\usetheme{Frankfurt}
\usecolortheme{crane}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{circuitikz}
\author{Maximilian Heim}
\title{Return Oriented Programming}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
\logo{\includegraphics[width=1.5cm]{1200px-Hsas_logo.svg.png}}
\institute{University Albstadt-Sigmaringen} 
\date{\today} 
\subject{Offensive Security Methods}
\begin{document}

@@

\section{Introduction}
\subsection{Return Oriented Programming}
\begin{frame}
    \frametitle{What is Return Oriented Programming?}
    \begin{enumerate}
    \item A type of buffer overflow attack
    \item Instead of writing executable code onto the stack we place return addresses onto the stack
    \item These get executed sequentially
    \item Using this technique we can have arbitrary code execution in large enough programs
    \end{enumerate}
\end{frame}
\subsection{Relevanz}
\begin{frame}
    \frametitle{Warum ist es wichtig diese zu erkennen?}
    \begin{itemize}
    \item Militär
    \item Finanzen
    \item Energie
    \item Geheimdienste
    \item Überwachungsstaaten
    \item Transport
    \item U.v.m \ldots
    \end{itemize}

@@

\section{Erkennung}
\subsection{Destruktive Erkennung}
\begin{frame}
    \frametitle{Reverse Engineering}
    \begin{itemize}
        \item Entfernen der Oberfläche Schicht für Schicht
        \item Advantages: 
        \begin{enumerate}
            \item 100 \% detection rate
        \end{enumerate}
        \item Disadvantages:
        \begin{enumerate}
            \item Only tests a single chip
            \item Chip unusable afterwards
            \item Very time consuming
        \end{enumerate}
        \item However, 
    \end{itemize}
\end{frame}
\subsection{Nicht-Destruktive Erkennung}
\section{Conclusion}
\begin{frame}
    \frametitle{Types of detection}
\end{frame}

>>>>

\documentclass[11pt]{beamer}
\usetheme{Frankfurt}
\usecolortheme{crane}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{fancyvrb}
%\usepackage{circuitikz}
\author{Maximilian Heim}
\title{Return Oriented Programming}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
\logo{\includegraphics[width=0.1\textwidth]{logo.png}}
\institute{University Albstadt-Sigmaringen} 
\date{\today} 
\subject{Offensive Security Methods}
\begin{document}

@@

\section{Introduction}
\subsection{Return Oriented Programming}
\begin{frame}
    \frametitle{What is Return Oriented Programming?}
    \begin{itemize}
    \item A type attack that exploits buffer overruns
    \item Published by Hovav Shacham in 2007
    \item Arised as a technique to counter security mechanisms (NX)
    \item Big binary $\rightarrow$ ROP is turing complete
    \item Many authors refer to ret-to-libc as ROP
    \end{itemize}
\end{frame}

\section{How does it work?}
\begin{frame}
    \frametitle{Overview}
    \begin{enumerate}
        \item Search the binary for gadgets: return (0xC3) bytes that contain useful instructions before
        \item Generate a list of these gadgets, called ROP chain
        \item Generate a payload with the addresses of these gadgets
        \item Insert payload via buffer overrun
    \end{enumerate}
\end{frame}
\begin{frame}
    \frametitle{ROP gadgets}
    \begin{itemize}
        \item Gadgets are machine instructions that end on a return
        \item Tools: ROPgadget (\url{https://github.com/JonathanSalwan/ROPgadget}), ropper (\url{https://github.com/sashs/Ropper}), Radare2, pwntools....
    \end{itemize}
    \begin{figure}[h]
        \caption{ROP Gadgets}
        \centering
        \includegraphics[width=0.8\textwidth]{gadget.png}\label{gadget}
    \end{figure}

@@

\begin{frame}[fragile]
    \frametitle{Useful gadgets: Write to register}
    \begin{itemize}
        \item Especially useful are pop instructions
    \end{itemize}
    \begin{Verbatim}
        POP eax, ret;
    \end{Verbatim}
    \begin{itemize}
        \item These allow us to write arbitrary values into registers
        \item However, sometimes we do not find a pop into our desired register (e.g. r14), here we can improvise and use something like
    \end{itemize}
    \begin{Verbatim}
        xor r14, r14
        pop r12
        xor r14, r12
    \end{Verbatim}
\end{frame}

\begin{frame}
    \frametitle{References}
    \url{https://trustfoundry.net/basic-rop-techniques-and-tricks/}
